---
title: "Reproducibility in Cancer Informatics"
---

```{r, include = FALSE}
leanbuild::set_knitr_image_path()
```
# Introduction

```{r, fig.alt="Title image: Reproducibility in Cancer Informatics", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gd422c5de97_0_0")
```

## Motivation

Cancer datasets are plentiful, complicated, and hold information that may be critical for the next research advancements. Cancer researchers may have analyzed data

This course introduces the concepts of reproducibility and replicability in the context of cancer informatics. It uses hands-on exercises to demonstrate in practical terms how to increase the reproducibility of data analyses. The course also introduces tools relevant to reproducibility including analysis notebooks, Docker images, git and GitHub.

## Target Audience  

The course is intended for students in the biomedical sciences and researchers who use informatics tools in their research.

## Curriculum  

```{r, fig.alt = "This course will demonstrate how to: Understanding why usability and documentation is vital, Identifying your user community, Building documentation and tutorials to maximize the usability of developed tools, Obtaining feedback from your users", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gd422c5de97_0_10")
```

The course includes a hands-on exercises for how to apply reproducible code concepts to their code
Individuals who take this course are encouraged to complete these activities as they follow along with the course material to help increase the reproducibility of their analyses.
