---
title: "Defining reproducibility"
---

```{r, include = FALSE}
leanbuild::set_knitr_image_path()
```

# Defining reproducibility

## Learning Objectives

```{r, fig.align='center', echo = FALSE, fig.alt= "This chapter will demonstrate how to: Understand the fundamental concepts of reproducibility vs replicability vs repeatability. Understand how working to make your analyses reproducible save your time and the time of your collaborators. Understand how working toward reproducible analyses enhances the accuracy of analysis results."}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf1accd298e_0_1205")
```

## What is reproducibility

There's been a lot of discussion about what is included in the term `reproducibility` and there is some discrepancy between fields. For the purposes of informatics and data analysis, a reproducible analysis is one that can be re-run by a different researcher and the same result and conclusion is found.

```{r, fig.align='center', echo = FALSE, fig.alt= "Reproducibility is a different analyst re­-performing the same analysis with the same code and data."}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf1accd298e_0_146")
```

Reproducibility is related to repeatability and replicability but it is worth taking the time parse out these terms.

```{r, fig.align='center', echo = FALSE, fig.alt= "Ruby the researcher has found something very interesting through data analysis. Ruby has a scatterplot on her computer that shows blue and pink data points and a trendline. The scatterplot has Variable A on the x axis and Variable B on the y axis. Ruby says my data analysis is showing a pattern that is very informative for ongoing research in my field."}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf1accd298e_0_0")
```


```{r, fig.align='center', echo = FALSE, fig.alt= "Repeatable means that if you keep everything the same but repeat the analysis - do you get the same results? Ruby the researcher has her same code and data but a repeat sign around them. If she re-runs the analysis, will she get the same scatterplot of results we’ve seen previously?"}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf1accd298e_0_413")
```

```{r, fig.align='center', echo = FALSE, fig.alt= "Ruby the researcher has found something very interesting through data analysis. Ruby has a scatterplot on her computer that shows blue data points and an trendline that shows a positive correlation. The scatterplot has Variable A on the x axis and Variable B on the y axis. Avi the associate sees Ruby the researcher’s results and is also excited about the findings. Avi says Ruby’s findings are super relevant to my work and I’m interested in using her methods!"}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf1accd298e_0_179")
```


```{r, fig.align='center', echo = FALSE, fig.alt= "Reproducible: using the same data and analysis but in the hands of another researcher - do we get the same results? Ruby the researcher sends her code and data to Avi the Associate who then tries to re-run the same analysis on his own computer."}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf1accd298e_0_368")
```


```{r, fig.align='center', echo = FALSE, fig.alt= "Replicable: with new data do we obtain the same inferences? Ruby the researcher sends her code to Avi the Associate who then tries to re-run the same analysis on his own computer but now with new data (represented by a different color flashdrive). Will Avi come to the same inference that Ruby’s analysis did?"}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf1accd298e_0_464")
```

## Reproducibility is critical

No doubt, reproducibility is critical to the scientific process moving along properly, but it also saves everyone involved time and effort!

```{r, fig.align='center', echo = FALSE, fig.alt= "Reproducible analyses save everyone time and effort!"}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf1cd772e00_0_5")
```
