---
title: "Using Notebooks"
---

```{r, include = FALSE}
leanbuild::set_knitr_image_path()
```

# Using Notebooks

## Learning Objectives

```{r, fig.align='center', echo = FALSE, fig.alt= "This chapter will demonstrate how to: Use an Integrated Development Environment to aid your development of code. Understand how notebooks can increase the usability and readability of scientific code. Set up code as a notebook. Use the notebook’s interface to interactively develop code."}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf8f405fdab_0_4")
```

## Readings about scientific computational notebooks

## Exercise: Convert code into a notebook!

#### Get the example project files for this chapter

For this chapter, we will create notebooks from our example files code!

<details> <summary> About command line </summary>

If you decide to follow the command line prompts but are not familiar with it [read this article](https://towardsdatascience.com/a-quick-guide-to-using-command-line-terminal-96815b97b955).
Make sure you have [`wget`](https://www.gnu.org/software/wget/manual/wget.html) installed on your computer. You can check if you have it by running the following in your command line:

```{bash}
wget -V
```

If this prints back something similar to this message above, then you already have `wget` and don't need to do anything! But if this command prints back `wget command not found` you will need to [ install wget](https://www.jcchouinard.com/wget/).

</details>

Now choose which version of the example you would like to work with and follow the instructions.  

<details> <summary>Python version of the exercise</summary>

To get the _Python project example files_, [click this link](https://raw.githubusercontent.com/jhudsl/Reproducibility-Examples/main/chapter-zips/python-heatmap-chapt-4.zip).
Or you can use these commands in command line.

```{bash, echo = FALSE, results = 'hide', message = FALSE, class.source = 'fold-show'}
mkdir -p chapter-zips
wget -O chapter-zips/python-heatmap-chapt-4.zip https://raw.githubusercontent.com/jhudsl/Reproducibility-Examples/main/chapter-zips/python-heatmap-chapt-4.zip
```

Now double click your chapter zip file to unzip. For Windows you may have to [follow these instructions](https://support.microsoft.com/en-us/windows/zip-and-unzip-files-f6dde0a7-0fec-8294-e1d3-703ed85e7ebc)).

```{bash, echo = FALSE, message = FALSE, results = 'hide', class.source = 'fold-show'}
unzip -o chapter-zips/python-heatmap-chapt-4.zip -d chapter-zips/
```

Now let's take a look at the files inside these projects. These are purposely not organized because for this exercise we will organize them!

```{bash, echo = FALSE, class.source = 'fold-show'}
ls chapter-zips/python-heatmap-chapt-4
```

Organize these files! For now we will organize these files by hand, but in the upcoming chapters we will make it so our analysis places these items in the correct directories (and creates the directories if they do not exist!).

- Create a `plots`, `results`, and `data` folder and organize the files into their respective folders. Note that `aggregated_metadata.json` and `LICENSE.TXT` also belong in the `data` folder.
- Delete any files that say "OLD". Keeping multiple versions of your scripts around is a recipe for mistakes and confusion. In the advanced course we will discuss how to use version control to help you track this more elegantly.

</details>

<details> <summary>R version of the exercise</summary>

To get the _R project examples files_ [click this link](https://raw.githubusercontent.com/jhudsl/Reproducibility-Examples/main/chapter-zips/r-heatmap-chapt-4.zip).
Or you can use these commands in command line.

```{bash, results = 'hide', message = FALSE, class.source = 'fold-show'}
mkdir -p chapter-zips
wget -O chapter-zips/r-heatmap-chapt-4.zip https://raw.githubusercontent.com/jhudsl/Reproducibility-Examples/main/chapter-zips/r-heatmap-chapt-4.zip
```

Now double click your chapter zip file to unzip. For Windows you may have to [follow these instructions](https://support.microsoft.com/en-us/windows/zip-and-unzip-files-f6dde0a7-0fec-8294-e1d3-703ed85e7ebc)).

```{bash, echo = FALSE, message = FALSE, results = 'hide', class.source = 'fold-show'}
unzip -o chapter-zips/r-heatmap-chapt-4.zip -d chapter-zips/
```

Now let's take a look at the files inside these projects. These are purposely not organized because for this exercise we will organize them!

```{bash, echo = FALSE, class.source = 'fold-show'}
ls chapter-zips/r-heatmap-chapt-4
```

Organize these files! For now we will organize these files by hand, but in the upcoming chapters we will make it so our analysis places these items in the correct directories (and creates the directories if they do not exist!).

- Create a `plots`, `results`, and `data` folder and organize the files into their respective folders. Note that `aggregated_metadata.json` and `LICENSE.TXT` also belong in the `data` folder.
- Delete any files that say "OLD". Keeping multiple versions of your scripts around is a recipe for mistakes and confusion. In the advanced course we will discuss how to use version control to help you track this more elegantly.

</details>

After your files are organized, you are ready to move on to the next chapter and start diving into the code!
