---
title: "Writing durable code"
---

```{r, include = FALSE}
leanbuild::set_knitr_image_path()
```

# Organizing your project

## Learning Objectives

```{r, fig.align='center', echo = FALSE, fig.alt= "This chapter will demonstrate how to: Identify what aspects make an analysis project more easily navigable. Set up a project with an organizational scheme that will work for the author and their colleagues."}
leanbuild::include_slide("https://docs.google.com/presentation/d/1LMurysUhCjZb7DVF6KS9QmJ5NBjwWVjRn40MS9f2noE/edit#slide=id.gf7bed24491_1_51")
```

## Get the exercise project files (or continue with the files you used in the previous chapter)

<details> <summary>How to get the _Python project example files_</summary>
To get the _Python project example files_, [click this link](https://raw.githubusercontent.com/jhudsl/Reproducibility_in_Cancer_Informatics/main/chapter-zips/python-heatmap-chapt-7.zip).

```{bash, include = FALSE}
mkdir -p chapter-zips
wget -O chapter-zips/python-heatmap-chapt-7.zip https://raw.githubusercontent.com/jhudsl/Reproducibility_in_Cancer_Informatics/main/chapter-zips/python-heatmap-chapt-7.zip
```

Now double click your chapter zip file to unzip. For Windows you may have to [follow these instructions](https://support.microsoft.com/en-us/windows/zip-and-unzip-files-f6dde0a7-0fec-8294-e1d3-703ed85e7ebc)).

```{bash, include = FALSE}
unzip -o chapter-zips/python-heatmap-chapt-7.zip -d chapter-zips/
```
</details>

<details> <summary>How to get the _R project example files_</summary>
To get the _R project example files_, [click this link](https://raw.githubusercontent.com/jhudsl/Reproducibility_in_Cancer_Informatics/main/chapter-zips/r-heatmap-chapt-7.zip).

```{bash, include = FALSE}
mkdir -p chapter-zips
wget -O chapter-zips/r-heatmap-chapt-7.zip https://raw.githubusercontent.com/jhudsl/Reproducibility_in_Cancer_Informatics/main/chapter-zips/r-heatmap-chapt-7.zip
```

Now double click your chapter zip file to unzip. For Windows you may have to [follow these instructions](https://support.microsoft.com/en-us/windows/zip-and-unzip-files-f6dde0a7-0fec-8294-e1d3-703ed85e7ebc)).

```{bash, include = FALSE}
unzip -o chapter-zips/r-heatmap-chapt-7.zip -d chapter-zips/
```
</details>

## Exercise: Make code more durable!
